---
title: "Practical Machine Learning - Assignment"
author: "Frantisek Mraz"
date: "Monday, August 25, 2014"
output: html_document
---


Alter loading the data using read.csv(), it appears a 
column named "X" which is simply index of the row. Hence I removed it and also 
the first real column containing user_name.
```{r}
library(caret)
training = read.csv("pml-training.csv")
training = training[c(-1,-2)]

testing = read.csv("pml-testing.csv")
testing = testing[c(-1,-2)]
```

The main problem with the data were undefined values NA and empty values. They 
can be removed

```{r}
isna = is.na(training)*1
trainNNA = training[colSums(isna)==0]

isnull = (trainNNA=="")*1
trainSmall = trainNNA[colSums(isnull)==0]
```

I have used random forest for fitting the data. In order to save time, I trained the model on a half of the available samples 

```{r}
train = sample(1:dim(trainSmall)[1],size=dim(trainSmall)[1]/2,replace=F)
trainSA = trainSmall[train,]
testSA = trainSmall[-train,]

set.seed(13234)

modelFit <- train(trainSA$classe ~ .,method="rf",data=trainSA)
print(modelFit$finalModel)
```

The model was good as it is seen from the confusion matrix for the test set `testSA`:

```{r}
predTest = predict(modelFit, testSA)
confusionMatrix(testSA$classe,predTest)
```

As the accuracy of the model is more than 99% on unseen data and the test data were so large (almost 10.000 samples), I did not made any cross-validation.

Before applying the model to other data, they must be preprocessed in the same way as the training data. For solving the assignment

```{r}
testNNA = testing[colSums(isna)==0]
testSmall = testNNA[colSums(isnull)==0]
predTestSmall = predict(modelFit, testSmall)
print(predTestSmall)
```
